


```{r}
# Load the data
the.fulldata <- as.matrix(read.csv("MelbAirportSolarData.csv", header = TRUE, sep = ","))

# Generate a sample of 5,000 data
my.data <- the.fulldata[sample(1:8928, 5000), ]

# Save the sample data to a text file
write.table(my.data, "RaviShankarJS-223296806-MelbAirSolarMyData.txt", row.names = FALSE, col.names = FALSE, sep = ",")
```

1.1) Histogram and box plot for the 'Wind speed' variable, and five-number summary:


```{r}
# Draw histogram
hist(my.data[, 3], main = "Histogram of Wind Speed", xlab = "Wind Speed (km/h)")

# Draw box plot
boxplot(my.data[, 3], main = "Box Plot of Wind Speed", ylab = "Wind Speed (km/h)")

# Five-number summary
wind_speed_summary <- summary(my.data[, 3])
print(wind_speed_summary)
```

1.2) Choosing summary statistics for center and spread of 'Wind speed':

We could use the following data to summarise the "Wind speed" variable's centre and spread:
Centre: The median, since it is a more reliable indicator of central tendency than the mean and is less impacted by outliers.
Spread: Interquartile range (IQR), as opposed to standard deviation, it offers a measure of the spread that is also resistant to outliers.
The IQR provides information about the typical range of the data, which is helpful in understanding the variability in the wind speed measurements, while the median provides the central value.

1.3) Scatterplot, linear regression, correlation, and coefficient of determination:

```{r}
# Scatterplot of Temperature vs. Humidity
plot(my.data[, 2], my.data[, 4], xlab = "Temperature (°C)", ylab = "Humidity (%)")

# Fit linear regression model
lm_model <- lm(my.data[, 4] ~ my.data[, 2])
abline(lm_model, col = "red")

# Print linear regression equation
print(summary(lm_model)$coefficients)

# Compute correlation coefficient and coefficient of determination
corr_coef <- cor(my.data[, 2], my.data[, 4])
r_squared <- summary(lm_model)$r.squared

cat("Linear regression equation:", lm_model$coefficients[1], "+", lm_model$coefficients[2], "* Temperature\n")
cat("Correlation coefficient:", corr_coef, "\n")
cat("Coefficient of determination (R-squared):", r_squared, "\n")
```

1.4) Constructing cross-tabulation and answering probability-related questions:

a) Construct the cross-tabulation:

```{r}
# Create new variables
WSB <- ifelse(my.data[, 3] > 25, "High", "Low")
TB <- cut(my.data[, 2], breaks = c(-Inf, 20, 30, Inf), labels = c("Low", "Moderate", "High"))
IrrB <- ifelse(my.data[, 1] > 800, "High", "Low")

# Construct cross-tabulation
cross_tab <- table(WSB, TB, IrrB)
print(cross_tab)
```

b) Use the above obtained cross table to answer the following questions. Show all the steps/workings clearly.
Consider that a record (row) is selected from the data at random,
i) what is the probability that the IrrB is High?

```{r}
# Calculate the total number of observations where IrrB is High
irr_high_count <- sum(cross_tab[, , "High"])

# Calculate the total number of observations
total_count <- sum(cross_tab)

# Calculate the probability that IrrB is High
prob_irr_high <- irr_high_count / total_count

cat("Probability that IrrB is High:", prob_irr_high, "\n")
```

ii) what is the probability that the TB is high given that the WSB is low?

```{r}
# Calculate the number of observations where TB is High and WSB is Low
tb_high_ws_low_count <- cross_tab["Low", "High", ]

# Calculate the total number of observations where WSB is Low
ws_low_count <- sum(cross_tab["Low", , ])

# Calculate the probability that TB is High given that WSB is Low
prob_tb_high_given_ws_low <- tb_high_ws_low_count / ws_low_count

cat("Probability that TB is High given that WSB is Low:", prob_tb_high_given_ws_low, "\n")
```

iii) what is the probability that the IrrB is low given that the TB is moderate and the WSB is low?

```{r}
# Calculate the number of observations where IrrB is Low, TB is Moderate, and WSB is Low
irr_low_tb_mod_ws_low_count <- cross_tab["Low", "Moderate", "Low"]

# Calculate the total number of observations where TB is Moderate and WSB is Low
tb_mod_ws_low_count <- sum(cross_tab["Low", "Moderate", ])

# Calculate the probability that IrrB is Low given that TB is Moderate and WSB is Low
prob_irr_low_given_tb_mod_ws_low <- irr_low_tb_mod_ws_low_count / tb_mod_ws_low_count

cat("Probability that IrrB is Low given that TB is Moderate and WSB is Low:", prob_irr_low_given_tb_mod_ws_low, "\n")
```


iv) Are high TB and low IrrB independent events? Explain.

```{r}
# Compute joint probabilities
joint_prob <- prop.table(cross_tab)

# Compute marginal probabilities
prob_tb_high <- sum(joint_prob[, "High", ])
prob_irr_low <- sum(joint_prob[, , "Low"])

# Compute expected joint probability under independence
expected_joint_prob <- prob_tb_high * prob_irr_low

# Compare observed and expected joint probabilities
observed_joint_prob <- joint_prob["High", "Low", "Low"]

if (abs(observed_joint_prob - expected_joint_prob) > 0.05) {
  cat("High TB and Low IrrB are not independent events.\n")
} else {
  cat("High TB and Low IrrB are independent events.\n")
}
```


v) Are low TB and high WSB mutually exclusive? Explain.

```{r}
# Check if there are any observations with both Low TB and High WSB
# Check if there are any observations with both Low TB and High WSB
if (sum(cross_tab["High", "Low", ]) == 0) {
  cat("Low TB and High WSB are mutually exclusive events.\n")
} else {
  cat("Low TB and High WSB are not mutually exclusive events.\n")
}
```

Q3)

Q3.2)

(C)

```{r}
# Load the required packages
library(ggplot2)
library(patchwork)

# Define the parameters
a <- 0.5  # Prior hyperparameter
b <- 10   # Prior hyperparameter
N <- 6    # Number of observations
x <- c(10, 12, 15, 20, 15, 20) # Observed dwell times

# Compute the posterior hyperparameters
a_posterior <- a + N
b_posterior <- b + sum(x)

# Create a grid of values for x (dwell time)
x_grid <- seq(min(x), max(x), length.out = 100)

# Calculate the likelihood, prior, and posterior distributions
likelihood <- dexp(x_grid, rate = 1/mean(x))
prior <- 1 / (dgamma(1/x_grid, shape = a, rate = 1/b))
posterior <- 1 / (dgamma(1/x_grid, shape = a_posterior, rate = 1/b_posterior))

# Create the plots
p1 <- ggplot() +
  geom_line(aes(x_grid, likelihood), color = "blue", linewidth = 1) +
  labs(x = "Dwell Time", y = "Density", title = "Likelihood Distribution")

p2 <- ggplot() +
  geom_line(aes(x_grid, prior), color = "red", linewidth = 1) +
  labs(x = "Dwell Time", y = "Density", title = "Prior Distribution")

p3 <- ggplot() +
  geom_line(aes(x_grid, posterior), color = "green", linewidth = 1) +
  labs(x = "Dwell Time", y = "Density", title = "Posterior Distribution")

# Combine the plots
(p1 | p2) / p3
```

Q4)

(C)

```{r}
# Define the prior distribution function
prior_distribution <- function(theta) {
  p <- 0
  if (theta >= 50 & theta <= 150) {
    p <- 1 / (45000 * (150 - 50))
  } else if (theta > 150 & theta <= 200) {
    p <- 1 / (90000 * (200 - 150))
  } else if (theta > 200 & theta <= 250) {
    p <- 1 / (30000 * (250 - 200))
  } else {
    p <- 0
  }
  return(p)
}

# Compute the posterior distribution
n <- 1
mean_prior <- 150
sigma_prior <- 20

# Likelihood function (normal distribution with known variance)
likelihood <- function(theta, y) {
  sigma <- 10
  return(dnorm(y, theta, sigma))
}

# Posterior distribution
posterior_distribution <- function(theta, y) {
  return(likelihood(theta, y) * prior_distribution(theta))
}

# Compute the posterior mean and standard deviation
y <- 160 # Observed mean height
theta_grid <- seq(50, 250, length.out = 1000)
posterior <- sapply(theta_grid, function(x) posterior_distribution(x, y))
posterior_mean <- sum(theta_grid * posterior) / sum(posterior)
posterior_var <- sum((theta_grid - posterior_mean)^2 * posterior) / sum(posterior)
posterior_sd <- sqrt(posterior_var)

cat("Posterior mean:", posterior_mean, "\n")
cat("Posterior standard deviation:", posterior_sd, "\n")

# Plot the prior distribution
prior_values <- sapply(theta_grid, prior_distribution)
plot(theta_grid, prior_values, type = "l", xlab = "theta", ylab = "Density", main = "Prior Distribution")

# Plot the likelihood function
likelihood_values <- sapply(theta_grid, function(x) likelihood(x, y))
plot(theta_grid, likelihood_values, type = "l", xlab = "theta", ylab = "Density", main = "Likelihood Function")

# Plot the posterior distribution
plot(theta_grid, posterior, type = "l", xlab = "theta", ylab = "Density", main = "Posterior Distribution")
```

Q5)

```{r}
# Load the data
zz <- read.table("lettersdata.txt")
zz <- as.matrix(zz)

# 5.1a) Draw a scatterplot and visually examine the number of clusters
plot(zz, pch = 20, xlab = "X", ylab = "Y")
# By visually examining the scatterplot, there appear to be around 4-5 clusters in the data.

# 5.1b) Perform k-means clustering with k = 4 (assuming 4 clusters from visual inspection)
set.seed(123)  # Set a seed for reproducibility
km_result <- kmeans(zz, centers = 4)
cluster_colors <- c("black", "red", "green", "blue")
plot(zz, col = cluster_colors[km_result$cluster], pch = 20, xlab = "X", ylab = "Y", main = "K-Means Clustering (k = 4)")
legend("topright", legend = paste("Cluster", 1:4), col = cluster_colors, pch = 20)
# The k-means algorithm has identified 4 clusters in the data, with some overlapping regions.

# 5.1c) Vary k from 2 to 20, record TOTWSS, and plot TOTWSS vs. k
set.seed(123)  # Set a seed for reproducibility
totwss <- numeric(19)
for (k in 2:20) {
  km_result <- kmeans(zz, centers = k)
  totwss[k - 1] <- km_result$tot.withinss
}
plot(2:20, totwss, type = "b", xlab = "Number of Clusters (k)", ylab = "Total Within Sum of Squares (TOTWSS)")
# The TOTWSS vs. k plot can be used to determine the optimal number of clusters by looking for an "elbow" or a point where the decrease in TOTWSS starts to level off. This typically indicates the appropriate number of clusters.

# 5.2) Perform spectral clustering with k = 4
library(kernlab)
sc_result <- specc(zz, centers = 4)
# Extract cluster assignments
sc_clusters <- sc_result@.Data
# Plot the results
plot(zz, col = sc_clusters + 1, pch = 20, xlab = "X", ylab = "Y", main = "Spectral Clustering (k = 4)")
legend("topright", legend = paste("Cluster", 1:4), col = 1:4, pch = 20)
# Spectral clustering also identifies 4 clusters in the data, but the cluster assignments may differ from k-means.
# Comparing the two plots, we can observe the differences in cluster assignments between k-means and spectral clustering.
```

